Charnley et al. 2021 - Exploring the relationships between drought and epidemic cholera in Africa 

```{r libraries}
library(dplyr)
library(magrittr)
library(corrplot)
library(ggplot2)
library(tidyr)
library(caret)
library(MASS)
```

Loading the data `data1` which consists of the outcome variable `Outbreaks` and all the environmental/socioeconomic data.

```{r load_data}
library(readxl)
data1 <- read_excel("Multivariate_model_data.xlsx", sheet = "Sheet2")
# remove missing entries and year/country column
data1$Year <- NULL
data1$Code <- NULL
data1 <- na.omit(data1)
```

Fetch the list of all covariates.

```{r covar_to_test}
#get full list of covariates
covar_to_test <- names(data1)[2:(ncol(data1))] # Starts at 2, to exclude outbreak

as.data.frame(covar_to_test)
```

Remove covariates that are not significantly associated with the data at p=<0.1, through fitting univariate models and examining the pvalues.

```{r fit_univariate}
mod <- list()
pvalues <- rep(NA, length(covar_to_test))
coeffs <- rep(NA, length(covar_to_test))
BICs <- rep(NA, length(covar_to_test))
for(i in 1:length(covar_to_test)){
  mod[[i]] <- glm( paste0("Outbreak~",covar_to_test[i]), 
           data = data1, family=binomial(link="cloglog"))
  
  pvalues[i] <- coef(summary(mod[[i]]))[,4]
  coeffs[i] <- coef(summary(mod[[i]]))[,1]
  BICs[i] <- BIC(mod[[i]])
}
df <-  data.frame(covariate = covar_to_test, 
                        pvalue = pvalues, 
                        coeff = coeffs,
                        BIC = BICs) %>% 
               arrange(-pvalues)
df
df %<>% filter(pvalue < 0.1)
covar_to_test = as.character(df$covariate)
```

We examine covariates that are highly correlated.

```{r remove_cor, fig.height=12, fig.width=12, fig.cap="Heatmap of parameter correlatons."}
reduced_dat <-  data1[, covar_to_test]
reduced_dat_cor <-  cor(reduced_dat)
heatmap(reduced_dat_cor, keep.dendro = TRUE)
cor_df <- as.data.frame(reduced_dat_cor) 
cor_df$param1 <- rownames(cor_df)
cor_df %<>% gather(param2, correlation, -param1)
```

\FloatBarrier

Group highly correlated covariates into clusters where the absolute piecewise correlation is above 0.75.

```{r clusters, fig.height=12, fig.width=10}
x <- as.dist( 1-abs(cor(na.omit(data1[, na.omit(covar_to_test)]))))
out <- hclust(x)
df <-  data.frame(param1 = unique(cor_df$param1))
i = length(covar_to_test)
pairwise_cor = 1
while(i >1 & pairwise_cor>0.75){ 
  i <- i-1
  
  y <- cutree(out, k = i)
  
  tmp = data.frame(param1 = names(y), as.numeric(y))
  names(tmp)[2] = paste0("cluster", i)
  
  df %<>% left_join(tmp)
  
  
  for(c in unique(y)){
    cluster_parm = df$param1[df[,paste0("cluster", i)] == c]
    if(length(cluster_parm)>1){
      out_cor = cor(as.matrix(data1[, as.character(cluster_parm)]))
      pairwise_cor = min(min(out_cor[out_cor<1]), pairwise_cor)
    }
  }
  
}
# final is i-1
df_out = df 
names(df_out)[2] = "cluster"
df_out %>% arrange(cluster)
```

Choose those covariates in each cluster that are most correlated with our outcome variable. 

```{r cor with cas}
df_out %<>% mutate(cor_with_case = NA)
for(i in 1:nrow(df_out)){
  df_out$cor_with_case[i] = abs(cor(data1[, c("Outbreak", 
                                            as.character(df_out$param1[i]))])[1,2])
}
df_covar <-  df_out %>% 
  group_by(cluster) %>% 
  arrange(-cor_with_case) %>% 
  summarise(param1 = first(param1))
covar_to_test = as.character(df_covar$param1)
covar_to_test
```

We use the `stepAIC` function from the `MASS` package to step through model possibilities. This uses BIC as a criteria to assess the inclusion of each model covariate.

```{r stepAIC}
dat_subs <-  data1[, c("Outbreak" ,covar_to_test)] # subset 
#use stepAIC to get list of important covariates
alt_mod <-  MASS::stepAIC(glm("Outbreak~.", data = dat_subs, 
                            family=binomial(link="cloglog")), 
                        trace = FALSE,
                        k = log(nrow(dat_subs))) #turning into BIC
summary(alt_mod)
# remove those that are not significant
pvalues <- coef(summary(alt_mod))[,4]
parm_out <- names(head(sort(pvalues), 16)) #get 15 most significant
parm_out <- parm_out[grep("(Intercept)", parm_out, invert = T)] # remove intercept
as.data.frame(parm_out)
```
